---
title: "Title"
author: "First Last"  
date: "last Updated: `r Sys.Date()`"
output:
  html_document:
    keep_md: yes
    toc: yes
    toc_float: yes
    code_folding: hide
    theme: flatly
    highlight: tango
editor_options: 
  chunk_output_type: console
---

This report (add a brief description and the purpose of the report)
Demonstration of the __COMBINED__ template

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of two directories.-->
```{r, echo=F, message=F, results = 'hide'}
cat("Working directory: ", getwd()) # turn on to test the location
# if the line above generates the project root, re-map by selecting 
# Session --> Set Working Directory --> To Source File location
# alternatively, set wd for knitting environment
library(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.
```


```{r set_options, echo=F}
# cat("Working directory: ", getwd()) # turn on to test the location
report_render_start_time <- Sys.time()
# set options shared by all chunks
knitr::opts_chunk$set(
  results      = 'show',
  message      = FALSE,
  warning      = FALSE,
  comment      = NA,
  tidy         = FALSE,
  # dpi        = 400, # dots per inch,
  # out.width  = "650px", # pixels, this affects only the markdown, not the underlying png file.  The height will be scaled appropriately.
  fig.width    = 9, # inches
  fig.height   = 4, # inches
  fig.path     = 'figure-png-com/' # where figures are stored
)
echo_chunks    <- FALSE #Toggle for debugging.
message_chunks <- FALSE #Toggle for debugging.
options(width=120) # number of characters to display in the output (dflt = 80)
ggplot2::theme_set(ggplot2::theme_bw()) # common theme for all graphs
```

# Environment

This section reviews the components of the working environment of the report. Non-technical readers are welcomed to skip. Come back if you need to understand the origins of custom functions, scripts, or data objects.

<details>
   <summary> Packages used <span class="glyphicon glyphicon-plus-sign"></span></summary>

Packages used in current report

<!-- Load packages, or at least verify they're available on the local machine.  Suppress the output when loading packages. -->
```{r load-packages, message=message_chunks, class.source = "fold-show"}
# Choose to be greedy: load only what's needed
# Three ways, from least (1) to most(3) greedy:
# -- 1.Attach these packages so their functions don't need to be qualified: http://r-pkgs.had.co.nz/namespace.html#search-path
library(ggplot2)   # graphs
library(forcats)   # factors
library(stringr)   # strings
library(lubridate) # dates
library(scales)    # formats
library(dplyr)     # manipulation
# -- 2.Import only certain functions of a package into the search path.
import::from("magrittr", "%>%")
# -- 3. Verify these packages are available on the machine, but their functions need to be qualified: http://r-pkgs.had.co.nz/namespace.html#search-path
requireNamespace("readr"    )# data import/export
requireNamespace("readxl"   )# data import/export
requireNamespace("tidyr"    )# tidy data
requireNamespace("janitor"  )# tidy data
requireNamespace("dplyr"    )# Avoid attaching dplyr, b/c its function names conflict with a lot of packages (esp base, stats, and plyr).
```
</details>


<details>
   <summary> External scripts <span class="glyphicon glyphicon-plus-sign"></span></summary>
```{r load-sources, class.source = "fold-show" }
source("../../scripts/common-functions.R")
```
</details>



<details>
   <summary> Global values <span class="glyphicon glyphicon-plus-sign"></span></summary>
Values used throughout the report.
```{r declare-globals, class.source = "fold-show"}
```
</details>



<details>
   <summary> Functions <span class="glyphicon glyphicon-plus-sign"></span></summary>
Custom functions defined for use in this report.
```{r declare-functions, message=message_chunks, echo=T, results="hide", class.source = "fold-show"}
```
</details>


# Data

To better understand the tables used in the report, you can click to expand a more technical description of the data origin. However, we suggest starting with the [What is a Row](#what-is-a-row) and [Data Description](#data-description) sections first. These sections provide a foundation for understanding the data and will help you better navigate the more technical description provided here. If you find that you still need more details, feel free to return to this section.


<details>
   <summary> Load data <span class="glyphicon glyphicon-plus-sign"></span></summary>

```{r load-data, results='show', message=FALSE,class.source = "fold-show"}

```
</details>


<details>
   <summary> Inspect data <span class="glyphicon glyphicon-plus-sign"></span></summary>
```{r inspect-data, results='show', message=FALSE, class.source = "fold-show", cache=TRUE}

```
</details>


<details>
   <summary> Inspect data <span class="glyphicon glyphicon-plus-sign"></span></summary>
```{r tweak-data-0, results='show', message=FALSE, class.source = "fold-show", cache=TRUE}

```
</details>


# What is a row {#what-is-a-row} 

Each row of `ds0` represents ...

```{r what-is-a-row }

```

# Data Description {#data-description}

```{r}

```


# Analysis

```{r table-1, echo = echo_chunks, results='show', message=message_chunks, class.source = "fold-show"}

```

```{r graph-1, echo=echo_chunks, fig.height=4, fig.width=9, message=FALSE, class.source=, results='show'}

```

```{r graph-2, echo = echo_chunks, results='show', message=message_chunks}

```

Session Information {#session-info}
===========================================================================

For the sake of documentation and reproducibility, the current report was rendered in the following environment.  Click the line below to expand.

<details>
  <summary>Environment <span class="glyphicon glyphicon-plus-sign"></span></summary>
```{r session-info, echo=FALSE}
if( requireNamespace("devtools", quietly = TRUE) ) {
  devtools::session_info()
} else {
  sessionInfo()
}
```
</details>

```{r session-duration, echo=FALSE}
report_render_duration_in_seconds <- round(as.numeric(difftime(Sys.time(), report_render_start_time, units="secs")))
```

Report rendered by `r Sys.info()["user"]` at `r strftime(Sys.time(), "%Y-%m-%d, %H:%M %z")` in `r report_render_duration_in_seconds` seconds.
