---
title: "Rmarkdown Annotation Layer"
author: "Andriy Koval"
date: "last Updated: `r Sys.Date()`"
output:
  html_document:
    keep_md: yes
    toc: yes
    toc_float: yes
    code_folding: show
    theme: simplex
    highlight: tango
    self-contained: true
editor: visual
editor_options: 
  chunk_output_type: inline
---

This report exemplifies a workflow in which executable script (source code layer) is coupled with the annotation layer to produce a dynamic document (e.g. html, pdf): `.R` + `.Rmd` = `.hmtl` files.

```{r echo=FALSE,results="hide",message=FALSE}
cat("Working directory: ", getwd())
library(knitr)
opts_knit$set(root.dir='../../') # keep this chunk isolated, don't combine 
```

```{r set_options, echo=F, results="hide", message=F}
# cat("Working directory: ", getwd()) # turn on to test the location
report_render_start_time <- Sys.time()
# set options shared by all chunks
opts_chunk$set(
  results      = 'show',
  message      = FALSE,
  warning      = FALSE,
  comment      = NA,
  tidy         = FALSE,
  # dpi        = 400, # dots per inch,
  # out.width  = "650px", # pixels, this affects only the markdown, not the underlying png file.  The height will be scaled appropriately.
  fig.width    = 6, # inches
  fig.height   = 4, # inches
  fig.path     = 'figure-png-iso/' # where figures from report are stored
)
echo_chunks    <- FALSE #Toggle for debugging.
message_chunks <- FALSE #Toggle for debugging.
options(width=100) # number of characters to display in the output (dflt = 80)
ggplot2::theme_set(ggplot2::theme_bw()) # common theme for all graphs
knitr::read_chunk("./analysis/report-example/source-code-layer.R") #This allows knitr to call chunks tagged in the underlying *.R file.
```

# Environment

<details>

<summary>Packages used</summary>

Packages used in current report <!-- Load packages, or at least verify they're available on the local machine.  Suppress the output when loading packages. -->

```{r load-packages, message=message_chunks, class.source = "fold-show"}
```

</details>

<details>

<summary>External scripts</summary>

<!-- Load the sources.  Suppress the output when loading sources. -->

```{r load-sources, message=message_chunks, class.source = "fold-show"}
```

</details>

<details>

<summary>Global values</summary>

Values used throughout the report. <!-- Load any Global functions and variables declared in the R file.  Suppress the output. -->

```{r declare-globals, message=message_chunks, class.source = "fold-show"}
```

</details>

<details>

<summary>Functions</summary>

Custom functions defined for use in this report.

```{r declare-functions, message=message_chunks, echo=T, results="hide",class.source = "fold-show"}
```

</details>

# Data

Consider a scenario in which rectangle `ds0` contains a set of survey responses

```{r load-data, warning=F}
```

What's inside this `analysis rectange`?

```{r inspect-data-0}
```

# Facts about the rectangle

Some facts about our analysis rectangle are good to know before any wrangling takes place.

## Q.1 What is the grain of this data set?

> Grain - the most granular level of the data, answers the question "What is a row?"

```{r}

ds0 %>% nrow()
ds0 %>% summarize(row_count=n())
ds0 %>% summarize(
  row_count = n()
  ,id_count = n_distinct(id)
) %>%
  mutate(
    id_is_the_grain = row_count == id_count
  )
```

## Q.2 What is the period of observation?

```{r}
ds0$date %>% summary() # looks like from [2011 to 2014]

ds0 %>% 
  summarize(
    earliest_date  = min(date,na.rm=T)
    ,latest_date   = max(date,na.rm=T)
    ,span_in_days  = (latest_date-earliest_date)
  )

```

# Transformations

Wrangle data to assist in analysis

```{r tweak-data-1}
```

```{r inspect-data-1}
ds0 %>% labelled::look_for() # compare the initial state to spot the diff
ds1 %>%  labelled::look_for()
# some summaries are more useful for various comparisons
explore::describe_all(ds0)
explore::describe_all(ds1)
# THe point is, must must provide a self-explanatory comparions of analysis rectangle BEFORE and AFTER a meaningful set of transformations. (e.g ds0 --> ds1)

ds1  %>%
  dplyr::distinct() %>%
  dplyr::arrange(desc(date)) # inspection of date variables
ds1 %>% tableone::CreateTableOne(data=.) %>% summary()
```

# Analysis

## Q.3 What is sex ratio?

```{r}
ds1 %>% group_by(sex) %>% count()
ds1 %>% group_by(sex) %>% summarize(row_count = n())
ds1 %>% group_by(sex) %>% summarize(respondent_count = n_distinct(id))

ds1 %>% 
  # mutate(weight = 1L) %>% 
  group_by(sex) %>% 
  summarize(
    respondent_count = n()
    ,client_count    = sum(weight)
  ) %>% 
  ungroup() %>% 
  mutate(
    respondent_share = respondent_count/sum(respondent_count)
    ,client_share    = client_count/sum(client_count)
)
# quick graph
ds1 %>% 
  group_by(sex) %>% 
  summarize(client_count = sum(weight)) %>%
  ggplot(aes(y = client_count, x = sex)) +
  geom_col()
```

Alternatively, organizing as sequence of objects (i.e. rectangle `ds1` begets plot `g1`) allows to be more versatile:

```{r}
g1 <-
  ds1 %>%
  group_by(sex) %>%
  summarize(client_count = sum(weight)) %>%
  ggplot(data=., aes(y = client_count, x = sex)) +
  geom_col()
g1 %>% print()
ggsave("../../analysis/report-example/g1.png",plot=g1)
ggsave("../../analysis/report-example/g1.png",plot=g1,width=6,height = 4) # optimize
ggsave("../../analysis/report-example/g1.jpg",plot=g1,width=21,height = 14,units = "cm") # optimize
```

## Q.4 Sex difference in the outcome

Is the outcome (employment) related to the sex of the client? Who is more likely to report being employed after the program?
