---
title: "Example: Combined in Rmarkdown"
author: "Andriy Koval"  
date: "last Updated: `r Sys.Date()`"
output:
  html_document:
    keep_md: yes
    toc: yes
    toc_float: yes
    code_folding: hide
    theme: flatly
    highlight: tango
    self-contained: true
editor_options: 
  chunk_output_type: console
---

This report exemplifies a workflow in which executable script (source code layer) and information displays for humans (annotation layer) are composed and managed in the a single `.Rmd` document, as opposed to a pair of linked `.R`-- `.Rmd` files.

```{r, echo=F, message=F, results = 'hide'}
cat("Working directory: ", getwd()) # turn on to test the location
# if the line above generates the project root, re-map by selecting 
# Session --> Set Working Directory --> To Source File location
# alternatively, set wd for knitting environment
library(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.
```

```{r set_options, echo=F}
# cat("Working directory: ", getwd()) # turn on to test the location
report_render_start_time <- Sys.time()
# set options shared by all chunks
knitr::opts_chunk$set(
  results      = 'show',
  message      = FALSE,
  warning      = FALSE,
  comment      = NA,
  tidy         = FALSE,
  # dpi        = 400, # dots per inch,
  # out.width  = "650px", # pixels, this affects only the markdown, not the underlying png file.  The height will be scaled appropriately.
  fig.width    = 9, # inches
  fig.height   = 4, # inches
  fig.path     = 'figure-png-com/' # where figures are stored
)
echo_chunks    <- FALSE #Toggle for debugging.
message_chunks <- FALSE #Toggle for debugging.
options(width=120) # number of characters to display in the output (dflt = 80)
ggplot2::theme_set(ggplot2::theme_bw()) # common theme for all graphs
```

# Environment

This section reviews the components of the working environment of the report. Non-technical readers are welcomed to skip. Come back if you need to understand the origins of custom functions, scripts, or data objects.

<details>

<summary>Packages used </summary>

Packages used in current report

<!-- Load packages, or at least verify they're available on the local machine.  Suppress the output when loading packages. -->

```{r load-packages, message=message_chunks, class.source = "fold-show"}
library(tidyverse)
```

</details>

<details>

<summary>External scripts </summary>

```{r load-sources, class.source = "fold-show" }
source("./scripts/common-functions.R")
source("./scripts/operational-functions.R")
```

</details>

<details>

<summary>Global values </summary>

Values used throughout the report.

```{r declare-globals, class.source = "fold-show"}
```

</details>

<details>

<summary>Functions </summary>

Custom functions defined for use in this report.

```{r declare-functions, message=message_chunks, echo=T, results="hide", class.source = "fold-show"}
```

</details>

# Data

To better understand the tables used in the report, you can click to expand a more technical description of the data origin. However, we suggest starting with the [What is a Row](#what-is-a-row) and [Data Description](#data-description) sections first. These sections provide a foundation for understanding the data and will help you better navigate the more technical description provided here. If you find that you still need more details, feel free to return to this section.

<details>

<summary>Load data </summary>

```{r load-data, results='show', message=FALSE,class.source = "fold-show"}
ds0 <- readr::read_rds("./data-public/raw/example-prosthetic-1.rds")
```

</details>

<details>

<summary>Inspect data </summary>

```{r inspect-data, results='show', message=FALSE, class.source = "fold-show", cache=TRUE}
ds0  %>% explore::describe_all()
ds0 %>% select(-id,-weight,-date) %>% tableone::CreateTableOne(data=.) %>% summary()
```

</details>

<details>

<summary>Transform data </summary>

```{r tweak-data-1, results='show', message=FALSE, class.source = "fold-show", cache=TRUE}
ds1 <-
  ds0 %>%
  rename(
    sex = gender # less biased term
  ) %>%
  mutate(
    male    = sex == "male"
    ,female = sex == "female"
    # ,caucasian  = race == "caucasian"
    # ,minority   = race == "minority"
    # ,aboriginal = race == "aboriginal"
    # ,employed_at_survey = employed == "employed"
  ) %>%
  mutate(
    year                = lubridate::year(date) %>% as.integer(),
    # yearmon             = tsibble::yearmonth, # not supported by look_for()
    # year_fiscal         = compute_fiscal_year(date),
    # quarter             = lubridate::quarter(date), # custom function, load from ./scripts/operational-functions.R
    # quarter_fiscal      = (quarter - 1),
    # quarter_fiscal      = ifelse(quarter_fiscal==0,4,quarter_fiscal)%>% as.integer(),
    # year_date           = as.Date(paste0(year,"-01-01")),
    # year_fiscal_date    = as.Date(paste0(year_fiscal,"-04-01")),
    # quarter_date        = paste(year,(quarter*3-1),"15", sep="-") %>% as.Date(),
    # quarter_fiscal_date = quarter_date,
  )
# compare the initial state to spot the diff
# some summaries are more useful for various comparisons
explore::describe_all(ds0)
explore::describe_all(ds1)
# THe point is, must must provide a self-explanatory comparions of analysis rectangle BEFORE and AFTER a meaningful set of transformations. (e.g ds0 --> ds1)
```

</details>

# What is a row? {#what-is-a-row}

## Q.1 What is the grain of this data set?

> Grain - the most granular level of the data, answers the question "What is a row?"

Each row of `ds0` represents a person

```{r what-is-a-row }
ds0 %>% nrow()
ds0 %>% summarize(row_count=n())
ds0 %>% summarize(
  row_count = n()
  ,id_count = n_distinct(id)
) %>%
  mutate(
    id_is_the_grain = row_count == id_count
  )
```

# Data Description {#data-description}

## Q.2 What is the period of observation?

```{r}
ds0$date %>% summary() # looks like from [2011 to 2014]

ds0 %>% 
  summarize(
    earliest_date  = min(date,na.rm=T)
    ,latest_date   = max(date,na.rm=T)
    ,span_in_days  = (latest_date-earliest_date)
  )
```

# Analysis

## Q.3 What is sex ratio?

### Table 1

```{r table-1, echo = echo_chunks, results='show', message=message_chunks, class.source = "fold-show"}
ds1 %>% group_by(sex) %>% count()
ds1 %>% group_by(sex) %>% summarize(row_count = n())
ds1 %>% group_by(sex) %>% summarize(respondent_count = n_distinct(id))

ds1 %>% 
  # mutate(weight = 1L) %>% # to test if weight changes the numbers
  group_by(sex) %>% 
  summarize(
    respondent_count = n()
    ,client_count    = sum(weight) # account for sampling and non-response weight
  ) %>% 
  ungroup() %>% 
  mutate(
    respondent_share = respondent_count/sum(respondent_count)
    ,client_share    = client_count/sum(client_count)
)
```

### Graph 1

We can create a graph on the fly...

```{r graph-1, echo=echo_chunks, fig.height=4, fig.width=9, message=FALSE, class.source='fold-hide', results='show'}
ds1 %>% group_by(sex) %>% summarize(client_count = sum(weight)) %>%
  ggplot(aes(y = client_count, x = sex)) +
  geom_col()
```

Alternatively, organizing as sequence of objects (i.e. rectangle `ds1` begets plot `g1`) allows to be more versatile:

```{r graph-2, echo = echo_chunks, results='show', message=message_chunks}
g1 <-
  ds1 %>%
  group_by(sex) %>%
  summarize(client_count = sum(weight)) %>%
  ggplot(data=., aes(y = client_count, x = sex)) +
  geom_col()
g1 %>% print()
ggsave("./analysis/report-example/g1.png",plot=g1)
ggsave("./analysis/report-example/g1.png",plot=g1,width=6,height = 4) # optimize
ggsave("./analysis/report-example/g1.jpg",plot=g1,width=21,height = 14,units = "cm") # optimize
```

# Session Information {#session-info}

For the sake of documentation and reproducibility, the current report was rendered in the following environment. Click the line below to expand.

<details>

<summary>Environment </summary>

```{r session-info, echo=FALSE}
if( requireNamespace("devtools", quietly = TRUE) ) {
  devtools::session_info()
} else {
  sessionInfo()
}
```

</details>

```{r session-duration, echo=FALSE}
report_render_duration_in_seconds <- round(as.numeric(difftime(Sys.time(), report_render_start_time, units="secs")))
```

Report rendered by `r Sys.info()["user"]` at `r strftime(Sys.time(), "%Y-%m-%d, %H:%M %z")` in `r report_render_duration_in_seconds` seconds.
