---
title: "Exploratory Data Analysis 1"
subtitle: "EDA 1"
author: "TBD"
format:
  html:
    page-layout: full
    toc: true
    toc-location: right
    code-fold: show
    theme: yeti
    highlight-style: breeze
    colde-line-numbers: true
    self-contained: true
    embed-resources: true
editor: visual
editor_options: 
  chunk_output_type: console
---

# Mission

This document introduces the reader to

# **Definition of Terms**

```{r}
#| code-summary: Reconciles working directory between execution modes
#| echo: false
#| results: hide
#| message: false 
#| warning: false
cat("Working directory: ", getwd())
library(knitr)
opts_knit$set(root.dir='../../') # keep this chunk isolated  
```

```{r}
#| label: set_options
#| code-summary: report-wide formatting options
#| echo: false
#| results: hide
#| message: false 
cat("Working directory: ", getwd()) # turn on to test the location
report_render_start_time <- Sys.time()
# set options shared by all chunks
opts_chunk$set(
  results      = 'show',
  message      = FALSE,
  warning      = FALSE,
  comment      = NA,
  tidy         = FALSE,
  # dpi        = 400, # dots per inch,
  out.width  = "960px", # pixels, this affects only the markdown, not the underlying png file.  The height will be scaled appropriately.
  fig.width    = 6, # inches
  fig.height   = 4, # inches
  fig.path     = 'figure-png-iso/' # where figures are stored
  
)
echo_chunks    <- FALSE #Toggle for debugging.
message_chunks <- FALSE #Toggle for debugging.
options(width=100) # number of characters to display in the output (dflt = 80)
ggplot2::theme_set(ggplot2::theme_bw()) # common theme for all graphs
read_chunk("./analysis/eda-1/eda-1.R") #This allows knitr to call
```

# Environment

> Reviews the components of the working environment of the report. Non-technical readers are welcomed to skip. Come back if you need to understand the origins of custom functions, scripts, or data objects.

```{r}
#| label: load-packages
#| code-summary: Packages used in the current report
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
```

```{r}
#| label: load-sources
#| code-summary: Collection of custom functions used in current repository
#| echo: true
#| results: hide
#| message: false
#| code-fold: true

```

```{r}
#| label: declare-globals
#| code-summary: Values used throughout the report.
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
```

```{r}
#| label: declare-functions
#| code-summary: Custom functions defined for use in this report.
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
```

```{r}
#| label: define-query
#| code-summary: Define queries to run
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
```

# Data

This report operates with the following data objects:

```{r }
#| label: load-data
#| code-summary: Load data objects 
#| warning: false
#| code-fold: true
#| cache: true
# set your root to Project Directory when developing chunks live in .qmd

```

## Tweak 0

```{r}
#| label: tweak-data-0
#| code-summary: Organize data into `ds0` form. Cleaning up names, optimize some data types.
#| code-fold: true
#| cache: true
#| results: hide

```

## Inspect starting files

```{r}
#| label: inspect-data-0
#| code-summary: Inspect contents of the available data tables
#| code-fold: false
#| eval: true
#| cache: true
ds_support %>% glimpse()
```

```{r}
#| label: inspect-data-1
#| code-summary: Inspect contents of the available data tables
#| code-fold: false
#| eval: true
#| cache: true
```

```{r}
#| label: inspect-data-2
#| code-summary: Inspect contents of the available data tables
#| code-fold: false
#| eval: true
#| cache: true
cat("QMD file has won")
```

## Tweak 1

```{r}
#| label: tweak-data-1
#| code-summary: Organize data into `ds0` form. Cleaning up names, optimize some data types.
#| code-fold: true
#| cache: true
#| results: hide
#| eval: false
```

# Fundamentals

```{r}
#| code-summary: Basic accounting 
#| code-fold: true
#| eval: true
#| cache: true

```