---
title: "Example: Combined in Quarto"
author: "Andriy Koval"
format:
  html:
    page-layout: full
    toc: true
    toc-location: right
    code-fold: show
    theme: simplex
    highlight-style: github
    colde-line-numbers: true
    css: styles.css
    self-contained: true
editor: visual
editor_options: 
  chunk_output_type: console
---

This report exemplifies a workflow in which executable script (source code layer) and information displays for humans (annotation layer) are composed and managed in the a single `.qmd` document, as opposed to a pair of linked `.R`-- `.qmd` files.

# Data

Let us assume that a rectangle `ds0` represents a set of survey responses for a given period of time.

```{r}
#| warning=F
library(tidyverse)
ds0 <- readr::read_rds("../../data-public/raw/example-prosthetic-2.rds")
```

What's inside this *analysis rectange*?

```{r}
# ds0 %>% View() #
# ds0 %>% glimpse()
ds0 %>% explore::describe_all()
# labelled::look_for(ds0)
# tableone::CreateTableOne(data = ds0)
ds0 %>% select(-id,-weight,-date) %>% tableone::CreateTableOne(data=.)
```

# Facts about the rectangle

Some facts about our analysis rectangle are good to know before any wrangling takes place.

## Q.1 What is the grain of this data set?

> Grain - the most granular level of data, answers the question "What is a row?"

```{r}
ds0 %>% nrow()
ds0 %>% summarize(row_count=n())
ds0 %>% summarize(
  row_count = n()
  ,id_count = n_distinct(id)
) %>%
  mutate(
    id_is_the_grain = row_count == id_count
  )
```

## Q.2 What is the period of observation?

```{r}
ds0$date %>% summary() # looks like from [2011 to 2014]

ds0 %>% 
  summarize(
    earliest_date  = min(date,na.rm=T)
    ,latest_date   = max(date,na.rm=T)
    ,span_in_days  = (latest_date-earliest_date)
  )

```

# Transformations

Wrangle data to assist in analysis

```{r}
ds1 <-
  ds0 %>%
  rename(
    sex = gender # less biased term
  ) %>%
  mutate(
    male    = sex == "male"
    ,female = sex == "female"
    # ,caucasian  = race == "caucasian"
    # ,minority   = race == "minority"
    # ,aboriginal = race == "aboriginal"
    # ,employed_at_survey = employed == "employed"
  ) %>%
  mutate(
    year                = lubridate::year(date) %>% as.integer(),
    # yearmon             = tsibble::yearmonth, # not supported by look_for()
    # year_fiscal         = compute_fiscal_year(date),
    # quarter             = lubridate::quarter(date), # custom function, load from ./scripts/operational-functions.R
    # quarter_fiscal      = (quarter - 1),
    # quarter_fiscal      = ifelse(quarter_fiscal==0,4,quarter_fiscal)%>% as.integer(),
    # year_date           = as.Date(paste0(year,"-01-01")),
    # year_fiscal_date    = as.Date(paste0(year_fiscal,"-04-01")),
    # quarter_date        = paste(year,(quarter*3-1),"15", sep="-") %>% as.Date(),
    # quarter_fiscal_date = quarter_date,
  )
ds0 %>% labelled::look_for() # compare the initial state to spot the diff
ds1 %>%  labelled::look_for()
# some summaries are more useful for various comparisons
explore::describe_all(ds0)
explore::describe_all(ds1)
# THe point is, must must provide a self-explanatory comparions of analysis rectangle BEFORE and AFTER a meaningful set of transformations. (e.g ds0 --> ds1)

ds1  %>%
  dplyr::distinct() %>%
  dplyr::arrange(desc(date)) # inspection of date variables
ds1 %>% tableone::CreateTableOne(data=.) %>% summary()
```

# Analysis

## Q.3 What is sex ratio?

```{r}
ds1 %>% group_by(sex) %>% count()
ds1 %>% group_by(sex) %>% summarize(row_count = n())
ds1 %>% group_by(sex) %>% summarize(respondent_count = n_distinct(id))

ds1 %>% 
  # mutate(weight = 1L) %>% 
  group_by(sex) %>% 
  summarize(
    respondent_count = n()
    ,client_count    = sum(weight)
  ) %>% 
  ungroup() %>% 
  mutate(
    respondent_share = respondent_count/sum(respondent_count)
    ,client_share    = client_count/sum(client_count)
)
# quick graph
ds1 %>% 
  group_by(sex) %>% 
  summarize(client_count = sum(weight)) %>%
  ggplot(aes(y = client_count, x = sex)) +
  geom_col()
```

Alternatively, organizing as sequence of objects (i.e. rectangle `ds1` begets plot `g1`) allows to be more versatile:

```{r}
g1 <-
  ds1 %>%
  group_by(sex) %>%
  summarize(client_count = sum(weight)) %>%
  ggplot(data=., aes(y = client_count, x = sex)) +
  geom_col()
g1 %>% print()
ggsave("../../analysis/report-example/g1.png",plot=g1)
ggsave("../../analysis/report-example/g1.png",plot=g1,width=6,height = 4) # optimize
ggsave("../../analysis/report-example/g1.jpg",plot=g1,width=21,height = 14,units = "cm") # optimize
```

## Q.4 Sex difference in the outcome

Is the outcome (employment) related to the sex of the client? Who is more likely to report being employed after the program?
