---
title: "Vehicle Performance Analysis: Exploratory Data Analysis"
subtitle: "EDA 1 - Motor Trend Car Road Tests Dataset"
author: "Analysis Team"
format:
  html:
    page-layout: full
    toc: true
    toc-location: right
    code-fold: show
    theme: yeti
    highlight-style: breeze
    code-line-numbers: true
    self-contained: true
    embed-resources: true
    keep-md: false
editor: visual
editor_options: 
  chunk_output_type: console
---

# Mission

This document demonstrates **data genealogy** and **graph families** within analytical workflows using the Motor Trend Car Road Tests dataset. We track data transformations through a clear lineage:

-   **ds0 (Original)**: Raw mtcars data with 32 individual vehicle observations
-   **g2_data (Derived)**: Cylinder-based performance summaries created in g2-data-prep chunk
-   **Graph Family g2x**: Multiple visualizations (g2, g21, g22...) that all build from the same g2_data ancestor
-   **Analytical Focus**: How engine cylinder count affects different aspects of vehicle performance

**Key Analytical Questions:**
 - How does cylinder count relate to horsepower? (g2 visualization) 
 - How does cylinder count relate to fuel efficiency? (g21 visualization) 
 - What other performance patterns emerge from the same cylinder-based data preparation? 
 - How does the graph family pattern help isolate data preparation while enabling multiple analytical perspectives?

Following the **template workflow**, we develop analytical insights iteratively in the companion `.R` script before publishing them here for broader consumption.

# **Definition of Terms**

```{r}
#| code-summary: Reconciles working directory between execution modes
#| echo: false
#| results: hide
#| message: false 
#| warning: false
cat("Working directory: ", getwd())
library(knitr)
opts_knit$set(root.dir='../../') # keep this chunk isolated
```

```{r}
#| label: set_options
#| code-summary: report-wide formatting options
#| echo: false
#| results: hide
#| message: false 
cat("Working directory: ", getwd()) # turn on to test the location
report_render_start_time <- Sys.time()
# set options shared by all chunks
opts_chunk$set(
  results      = 'show',
  message      = FALSE,
  warning      = FALSE,
  comment      = NA,
  tidy         = FALSE,
  # dpi        = 400, # dots per inch,
  out.width  = "960px", # pixels, this affects only the markdown, not the underlying png file.  The height will be scaled appropriately.
  fig.width    = 10, # inches - larger for publication trends
  fig.height   = 6, # inches - good for time series
  fig.path     = 'figure-png-iso/' # where figures are stored
  
)
echo_chunks    <- FALSE #Toggle for debugging.
message_chunks <- FALSE #Toggle for debugging.
options(width=100) # number of characters to display in the output (dflt = 80)
ggplot2::theme_set(ggplot2::theme_minimal()) # clean theme for publications
read_chunk("analysis/eda-1/eda-1.R") #This allows knitr to call chunks from R script
```

# Environment

> Reviews the components of the working environment of the report. Non-technical readers are welcomed to skip. Come back if you need to understand the origins of custom functions, scripts, or data objects.

```{r}
#| label: load-packages
#| code-summary: Packages used in the current report
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
```

```{r}
#| label: load-sources
#| code-summary: Collection of custom functions used in current repository
#| echo: true
#| results: hide
#| message: false
#| code-fold: true

```

```{r}
#| label: declare-globals
#| code-summary: Values used throughout the report.
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
```

```{r}
#| label: declare-functions
#| code-summary: Custom functions defined for use in this report.
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
```

# Data

This report demonstrates **data genealogy** and **graph families** using the Motor Trend Car Road Tests dataset:

-   **ds0**: Original `mtcars` data - 32 automobile performance measurements from 1973-74 Motor Trend magazine
-   **g2_data**: Derived cylinder summary created in `g2-data-prep` chunk - aggregated performance metrics by engine cylinder count (4, 6, 8 cylinders)

**Graph Family Pattern:** 
- ds0 → g2-data-prep → g2_data → **multiple visualizations** (g2, g21, g22...) 
- **Key insight**: All family members share the same data preparation but explore different aspects of monothematic idea (e.g  cylinder-based performance)

```{r }
#| label: load-data
#| code-summary: Load original data (ds0) - g2_data will be created in g2-data-prep chunk
#| warning: false
#| code-fold: true
#| cache: true
```

```{r}
#| label: inspect-data-0
#| code-summary: Data overview - ds0 structure
#| code-fold: false
#| eval: true
#| cache: true
```

```{r}
#| label: inspect-data-1
#| code-summary: Original dataset structure (ds0)
#| code-fold: false
#| eval: true
#| cache: true
```

```{r}
#| label: inspect-data-2
#| code-summary: Key variables summary from ds0
#| code-fold: false
#| eval: true
#| cache: true
```

# Analysis

## Data Genealogy Visualization

### Individual Vehicle Analysis (ds0)

This visualization uses our **original dataset (ds0)** to explore relationships between individual vehicle characteristics:

```{r}
#| label: g1
#| code-summary: Weight vs MPG scatter plot using original data (ds0)
#| code-fold: true
#| eval: true
#| cache: true
#| fig-cap: "Scatter plot of vehicle weight vs fuel efficiency - original data (ds0) with individual observations"
```

### Data Preparation for Aggregated Analysis

First, we prepare the data for our cylinder analysis:

```{r}
#| label: g2-data-prep
#| code-summary: Prepare g2_data - cylinder-based performance summary
#| code-fold: true
#| eval: true
#| cache: true
```

### Aggregated Performance Analysis (g2_data)

This visualization uses our **derived dataset (g2_data)** created above to show summarized performance metrics:

```{r}
#| label: g2
#| code-summary: Average horsepower by cylinder count using g2_data
#| code-fold: true
#| eval: true
#| cache: true
#| fig-cap: "Bar chart showing average horsepower by engine cylinder count - g2_data aggregation"
```

## Graph Family g2x: Different Perspectives on Cylinder Performance

### Alternative View: Fuel Efficiency Focus (g21)

This visualization uses the **same g2_data** as the previous chart, but focuses on fuel efficiency instead of horsepower. This demonstrates the graph family concept - same data preparation, different analytical perspectives:

```{r}
#| label: g21
#| code-summary: Average fuel efficiency by cylinder count using g2_data
#| code-fold: true
#| eval: true
#| cache: true
#| fig-cap: "Fuel efficiency by cylinder count - same g2_data ancestor, different performance metric focus"
print(g21_cylinder_mpg)  # Print for HTML rendering in Quarto
```

<!-- As we develop more graphs in the R script, we'll add corresponding chunks here -->

````{=html}
<!-- Example pattern:
### Language Dynamics

```{r}
#| label: g2
#| code-summary: Ukrainian vs Russian language publication trends
#| code-fold: false
#| eval: true
#| cache: true
#| fig-cap: "Publishing trends by language: Ukrainian and Russian books over time"
```

### Regional Patterns

```{r}
#| label: g3
#| code-summary: Regional distribution of publishing activity
#| code-fold: false
#| eval: true
#| cache: true
#| fig-cap: "Publishing activity by Ukrainian oblast/territory"
```
-->
````

# Session Info

```{r}
#| label: session-info
#| code-summary: R session information for reproducibility
#| echo: false
#| code-fold: true
cat("Report completed at:", as.character(Sys.time()), "\n")
cat("Total runtime:", round(difftime(Sys.time(), report_render_start_time, units = "mins"), 2), "minutes\n")
sessionInfo()
```