---
title: "Publishing Trends in Ukraine: Exploratory Data Analysis"
subtitle: "EDA 3 - Books of Ukraine Dataset"
author: "Research Group FIDES"
format:
  html:
    page-layout: full
    toc: true
    toc-location: right
    code-fold: show
    theme: yeti
    highlight-style: breeze
    code-line-numbers: true
    self-contained: true
    embed-resources: true
    keep-md: true
editor: visual
editor_options: 
  chunk_output_type: console
---

# Mission

This document explores publishing trends in Ukraine since 2005, focusing on:

-   **Temporal patterns**: How publication volumes have changed over time
-   **Language dynamics**: Ukrainian vs Russian language publishing trends\
-   **Regional differences**: Geographic variations in publishing activity
-   **Genre patterns**: Subject matter evolution across time and regions

Following the **FIDES framework**, we develop analytical insights iteratively in the companion `.R` script before publishing them here for broader consumption.

# **Definition of Terms**

```{r}
#| code-summary: Reconciles working directory between execution modes
#| echo: false
#| results: hide
#| message: false 
#| warning: false
cat("Working directory: ", getwd())
library(knitr)
opts_knit$set(root.dir='../../') # keep this chunk isolated
```

```{r}
#| label: set_options
#| code-summary: report-wide formatting options
#| echo: false
#| results: hide
#| message: false 
cat("Working directory: ", getwd()) # turn on to test the location
report_render_start_time <- Sys.time()
# set options shared by all chunks
opts_chunk$set(
  results      = 'show',
  message      = FALSE,
  warning      = FALSE,
  comment      = NA,
  tidy         = FALSE,
  # dpi        = 400, # dots per inch,
  out.width  = "960px", # pixels, this affects only the markdown, not the underlying png file.  The height will be scaled appropriately.
  fig.width    = 10, # inches - larger for publication trends
  fig.height   = 6, # inches - good for time series
  fig.path     = 'figure-png-iso/' # where figures are stored
  
)
echo_chunks    <- FALSE #Toggle for debugging.
message_chunks <- FALSE #Toggle for debugging.
options(width=100) # number of characters to display in the output (dflt = 80)
ggplot2::theme_set(ggplot2::theme_minimal()) # clean theme for publications
read_chunk("analysis/eda-1/eda-1.R") #This allows knitr to call chunks from R script
```

# Environment

> Reviews the components of the working environment of the report. Non-technical readers are welcomed to skip. Come back if you need to understand the origins of custom functions, scripts, or data objects.

```{r}
#| label: load-packages
#| code-summary: Packages used in the current report
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
```

```{r}
#| label: load-sources
#| code-summary: Collection of custom functions used in current repository
#| echo: true
#| results: hide
#| message: false
#| code-fold: true

```

```{r}
#| label: declare-globals
#| code-summary: Values used throughout the report.
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
```

```{r}
#| label: declare-functions
#| code-summary: Custom functions defined for use in this report.
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
```

# Data

This report operates with the **Books of Ukraine analytical database** - the final, analysis-ready dataset created by the Ellis pipeline.

```{r }
#| label: load-data
#| code-summary: Load data objects from Books of Ukraine database
#| warning: false
#| code-fold: true
#| cache: true
```

```{r}
#| label: inspect-data-0
#| code-summary: Inspect loaded data tables structure
#| code-fold: false
#| eval: true
#| cache: true
# Summary of available tables will be shown here
```

```{r}
#| label: inspect-data-1
#| code-summary: Examine data dimensions and key variables
#| code-fold: false
#| eval: true
#| cache: true
```

```{r}
#| label: inspect-data-2
#| code-summary: Quick data quality checks
#| code-fold: false
#| eval: true
#| cache: true
```

# Analysis

## Publishing Volume Trends

### Overall Publication Patterns

```{r}
#| label: g1
#| code-summary: Annual publication volume trends
#| code-fold: true
#| eval: true
#| cache: true
#| fig-cap: "Total number of unique books published in Ukraine by year (2005-2023)"
```

<!-- As we develop more graphs in the R script, we'll add corresponding chunks here -->

````{=html}
<!-- Example pattern:
### Language Dynamics

```{r}
#| label: g2
#| code-summary: Ukrainian vs Russian language publication trends
#| code-fold: false
#| eval: true
#| cache: true
#| fig-cap: "Publishing trends by language: Ukrainian and Russian books over time"
```

### Regional Patterns

```{r}
#| label: g3
#| code-summary: Regional distribution of publishing activity
#| code-fold: false
#| eval: true
#| cache: true
#| fig-cap: "Publishing activity by Ukrainian oblast/territory"
```
-->
````

# Session Info

```{r}
#| label: session-info
#| code-summary: R session information for reproducibility
#| echo: false
#| code-fold: true
cat("Report completed at:", as.character(Sys.time()), "\n")
cat("Total runtime:", round(difftime(Sys.time(), report_render_start_time, units = "mins"), 2), "minutes\n")
sessionInfo()
```