---
title: "{TITLE}"
subtitle: "{NAME} — {TYPE} Report"
author: "Analysis Team"
date: last-modified
format:
  html:
    page-layout: full
    toc: true
    toc-location: right
    code-fold: show
    theme: yeti
    highlight-style: breeze
    code-line-numbers: true
    self-contained: true
    embed-resources: true
    keep-md: false
editor: visual
editor_options:
  chunk_output_type: console
---

<!-- Composing Orchestra template — customize all {PLACEHOLDERS} before use. -->
<!-- Mode: {TYPE} (EDA = explore with open mind | Report = synthesize prior findings) -->

# Mission

{1–2 sentence mission from report-contract.prompt.md}

**Research Questions:**

1. {RQ1}
2. {RQ2}
3. {RQ3}

Following the dual-file development pattern, analytical code lives in the companion `.R` script. This document sources those chunks for presentation.

# Definition of Terms

```{r}
#| code-summary: Reconciles working directory between execution modes
#| echo: false
#| results: hide
#| message: false
#| warning: false
cat("Working directory: ", getwd())
library(knitr)
opts_knit$set(root.dir='../../') # keep this chunk isolated
```

```{r}
#| label: set_options
#| code-summary: Report-wide formatting options
#| echo: false
#| results: hide
#| message: false
cat("Working directory: ", getwd())
report_render_start_time <- Sys.time()
opts_chunk$set(
  results      = 'show',
  message      = FALSE,
  warning      = FALSE,
  comment      = NA,
  tidy         = FALSE,
  out.width    = "960px",
  fig.width    = 10,
  fig.height   = 6,
  fig.path     = 'figure-png-iso/'
)
echo_chunks    <- FALSE
message_chunks <- FALSE
options(width = 100)
ggplot2::theme_set(ggplot2::theme_minimal())
read_chunk("analysis/{NAME}/{NAME}.R")
```

# Environment

> Reviews the components of the working environment. Non-technical readers may skip this section.

```{r}
#| label: load-packages
#| code-summary: Packages used in the current report
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
```

```{r}
#| label: load-sources
#| code-summary: Collection of custom functions used in current repository
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
```

```{r}
#| label: declare-globals
#| code-summary: Values used throughout the report
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
```

```{r}
#| label: declare-functions
#| code-summary: Custom functions defined for use in this report
#| echo: true
#| results: hide
#| message: false
#| code-fold: true
```

# Data

<!-- Describe the data assets used in this report. Reference the data primer for details. -->

```{r}
#| label: load-data
#| code-summary: Load data from Ellis parquet outputs
#| warning: false
#| code-fold: true
#| cache: true
```

```{r}
#| label: inspect-data-0
#| code-summary: Data overview
#| code-fold: false
#| eval: true
#| cache: true
```

```{r}
#| label: inspect-data-1
#| code-summary: Grain verification
#| code-fold: false
#| eval: true
#| cache: true
```

# Analysis

## Graph Family g1: {g1 description from contract}

<!-- Research question: {RQ1} -->

```{r}
#| label: g1-data-prep
#| code-summary: Prepare data for g1 family
#| code-fold: true
#| eval: true
#| cache: true
```

```{r}
#| label: g1
#| code-summary: "{g1 brief description}"
#| code-fold: true
#| eval: true
#| cache: true
#| fig-cap: "{g1 caption}"
```

## Graph Family g2: {g2 description from contract}

<!-- Research question: {RQ2} -->

```{r}
#| label: g2-data-prep
#| code-summary: Prepare data for g2 family
#| code-fold: true
#| eval: true
#| cache: true
```

```{r}
#| label: g2
#| code-summary: "{g2 brief description}"
#| code-fold: true
#| eval: true
#| cache: true
#| fig-cap: "{g2 caption}"
```

## Graph Family g3: {g3 description from contract}

<!-- Research question: {RQ3} -->

```{r}
#| label: g3-data-prep
#| code-summary: Prepare data for g3 family
#| code-fold: true
#| eval: true
#| cache: true
```

```{r}
#| label: g3
#| code-summary: "{g3 brief description}"
#| code-fold: true
#| eval: true
#| cache: true
#| fig-cap: "{g3 caption}"
```

# Session Information

```{r}
#| code-summary: Session info and render time
#| echo: false
cat("Report rendered in", round(difftime(Sys.time(), report_render_start_time, units = "secs"), 1), "seconds\n")
sessionInfo()
```
