---
title: "Data Primer — {NAME}"
subtitle: "Canonical data documentation: Ellis outputs, schemas, variable reference"
author: "Report Composer (Composing Orchestra)"
date: last-modified
format:
  html:
    page-layout: full
    toc: true
    toc-location: right
    code-fold: show
    theme: yeti
    highlight-style: breeze
    code-line-numbers: true
    self-contained: true
    embed-resources: true
editor: visual
editor_options:
  chunk_output_type: console
---

<!-- Composing Orchestra data primer template.
     Used for analysis/data-primer-1/ — the CENTRALIZED data reference for the project.
     Composed once via @report-composer, verified by humans, linked by all EDAs and Reports.
     Customize {PLACEHOLDERS} and populate sections from Ellis metadata + CACHE manifest. -->

# Purpose

This is the **canonical data primer** for the project. It provides the foundational documentation of all data assets produced by the Ellis lane. Every EDA and presentation Report in `analysis/` links here for data context.

This primer answers: *What data do we have? What does it look like? What do the variables mean?*

For domain terminology, see `ai/project/glossary.md`.

```{r}
#| code-summary: Reconciles working directory
#| echo: false
#| results: hide
#| message: false
#| warning: false
cat("Working directory: ", getwd())
library(knitr)
opts_knit$set(root.dir='../../')
```

```{r}
#| label: setup
#| code-summary: Environment setup
#| echo: false
#| results: hide
#| message: false
library(magrittr)
library(dplyr)
library(tidyr)
library(arrow)
library(knitr)
library(scales)

opts_chunk$set(
  results   = 'show',
  message   = FALSE,
  warning   = FALSE,
  comment   = NA,
  tidy      = FALSE,
  out.width = "960px"
)
options(width = 100)
```

# Pipeline Architecture

The data assets used in this project follow a three-stage provenance chain:

```
Batch-91 source data (M:/.../batch-91/)
  → Ellis lane (manipulation/0-ellis.R)
    → R-friendly parquet files (data-private/derived/manipulation/)
```

**Ellis** imports the batch-91 CACHE tables, applies R-friendly type conversions (factors for categoricals, dates, integers for IDs), validates against the CACHE manifest, and outputs optimized parquet files.

# Data Assets Inventory

```{r}
#| label: inventory
#| code-summary: List available parquet files from Ellis output
#| code-fold: true

# List Ellis outputs
ellis_path <- "./data-private/derived/manipulation/"
if (dir.exists(ellis_path)) {
  parquet_files <- list.files(ellis_path, pattern = "\\.parquet$", full.names = TRUE)
  inventory <- data.frame(
    File = basename(parquet_files),
    stringsAsFactors = FALSE
  )
  # Read schema info for each file
  inventory$Rows <- NA_integer_
  inventory$Columns <- NA_integer_
  for (i in seq_along(parquet_files)) {
    tryCatch({
      ds_temp <- arrow::read_parquet(parquet_files[i])
      inventory$Rows[i] <- nrow(ds_temp)
      inventory$Columns[i] <- ncol(ds_temp)
    }, error = function(e) {
      message("Could not read: ", parquet_files[i])
    })
  }
  kable(inventory, caption = "Ellis Output: Analysis-Ready Parquet Files")
} else {
  cat("Ellis output directory not found at:", ellis_path, "\n")
  cat("Run the Ellis lane first: manipulation/0-ellis.R\n")
}
```

# Table Schemas

## Client Roster

```{r}
#| label: schema-client-roster
#| code-summary: Client roster structure
#| code-fold: true

path_client <- paste0(ellis_path, "client_roster.parquet")
if (file.exists(path_client)) {
  ds_client <- arrow::read_parquet(path_client)
  cat("Grain: person_oid (one row per individual)\n")
  cat("Rows:", scales::comma(nrow(ds_client)), "| Columns:", ncol(ds_client), "\n\n")
  glimpse(ds_client)
} else {
  cat("File not found:", path_client, "\n")
}
```

## Support by Year (Financial Support)

```{r}
#| label: schema-support
#| code-summary: Support by year structure
#| code-fold: true

path_support <- paste0(ellis_path, "support_by_year.parquet")
if (file.exists(path_support)) {
  ds_support <- arrow::read_parquet(path_support)
  cat("Grain: person_oid × year (one row per person per year of financial support)\n")
  cat("Rows:", scales::comma(nrow(ds_support)), "| Columns:", ncol(ds_support), "\n\n")
  glimpse(ds_support)
} else {
  cat("File not found:", path_support, "\n")
}
```

## Training by Year

```{r}
#| label: schema-training
#| code-summary: Training by year structure
#| code-fold: true

path_training <- paste0(ellis_path, "training_by_year.parquet")
if (file.exists(path_training)) {
  ds_training <- arrow::read_parquet(path_training)
  cat("Grain: person_oid × year (one row per person per year of training services)\n")
  cat("Rows:", scales::comma(nrow(ds_training)), "| Columns:", ncol(ds_training), "\n\n")
  glimpse(ds_training)
} else {
  cat("File not found:", path_training, "\n")
}
```

## Assessment by Year

```{r}
#| label: schema-assessment
#| code-summary: Assessment by year structure
#| code-fold: true

path_assessment <- paste0(ellis_path, "assessment_by_year.parquet")
if (file.exists(path_assessment)) {
  ds_assessment <- arrow::read_parquet(path_assessment)
  cat("Grain: person_oid × assessment event\n")
  cat("Rows:", scales::comma(nrow(ds_assessment)), "| Columns:", ncol(ds_assessment), "\n\n")
  glimpse(ds_assessment)
} else {
  cat("File not found:", path_assessment, "\n")
}
```

# Grain Demonstration

```{r}
#| label: grain-check
#| code-summary: Verify the grain of each table
#| code-fold: true

if (exists("ds_client") && is.data.frame(ds_client)) {
  cat("--- Client Roster ---\n")
  n_total <- nrow(ds_client)
  n_unique <- ds_client %>% distinct(person_oid) %>% nrow()
  cat("Total rows:", scales::comma(n_total), "\n")
  cat("Unique person_oid:", scales::comma(n_unique), "\n")
  cat("Grain holds (1:1):", n_total == n_unique, "\n\n")
}

if (exists("ds_support") && is.data.frame(ds_support)) {
  cat("--- Support by Year ---\n")
  n_total <- nrow(ds_support)
  n_unique <- ds_support %>% distinct(person_oid, year) %>% nrow()
  cat("Total rows:", scales::comma(n_total), "\n")
  cat("Unique person_oid × year:", scales::comma(n_unique), "\n")
  cat("Grain holds (1:1):", n_total == n_unique, "\n\n")
}
```

# Single Person View

```{r}
#| label: single-person
#| code-summary: What the data looks like for one individual
#| code-fold: true

# Select one person who appears in multiple tables for a rich example
if (exists("ds_support") && is.data.frame(ds_support)) {
  example_oid <- ds_support %>%
    count(person_oid) %>%
    filter(n >= 3, n <= 10) %>%  # someone with moderate history
    slice_sample(n = 1) %>%
    pull(person_oid)

  cat("=== Example person_oid:", example_oid, "===\n\n")

  if (exists("ds_client") && is.data.frame(ds_client)) {
    cat("--- Client Roster record ---\n")
    ds_client %>% filter(person_oid == example_oid) %>% glimpse()
    cat("\n")
  }

  cat("--- Financial Support history ---\n")
  ds_support %>% filter(person_oid == example_oid) %>% print()
  cat("\n")

  if (exists("ds_training") && is.data.frame(ds_training)) {
    training_records <- ds_training %>% filter(person_oid == example_oid)
    if (nrow(training_records) > 0) {
      cat("--- Training history ---\n")
      print(training_records)
      cat("\n")
    } else {
      cat("--- No training records for this individual ---\n\n")
    }
  }

  if (exists("ds_assessment") && is.data.frame(ds_assessment)) {
    assessment_records <- ds_assessment %>% filter(person_oid == example_oid)
    if (nrow(assessment_records) > 0) {
      cat("--- Assessment history ---\n")
      print(assessment_records)
    } else {
      cat("--- No assessment records for this individual ---\n")
    }
  }
} else {
  cat("Support data not loaded. Run Ellis lane first.\n")
}
```

# Variable Reference

Key variables across Ellis tables, sourced from `data-public/metadata/CACHE-manifest.md`:

## Identifiers
| Variable | Type | Description |
|----------|------|-------------|
| `person_oid` | integer | Unique person identifier across all tables |
| `year` | integer | Calendar year of the service event |

## Demographics (available in support, training, assessment tables)
| Variable | Type | Levels | Notes |
|----------|------|--------|-------|
| `age_category3` | factor | 3 levels | Age bracket (coarse) |
| `sex2` | factor | 2 levels | Male/Female |
| `marital2` | factor | 2 levels | Married/Not married |
| `disability2` | factor | 2 levels | Yes/No |
| `dependent2` | factor | 2 levels | Has dependents: Yes/No |
| `education3` | factor | 3 levels | Education level (coarse) |
| `ethnicity` | factor | varies | Ethnic background |
| `immigration` | factor | varies | Immigration status |

## Financial Support Specific
| Variable | Type | Description |
|----------|------|-------------|
| `client_type_code` | integer | Program classification code (see glossary) |
| `program_class0` | factor | FS (Financial Support) |
| `program_class1` | factor | IS / OTI / AISH / DRES |
| `program_class2` | factor | ETW / BFE / OTI / AISH |
| `program_class3` | factor | Detailed program description |

## Program Taxonomy Quick Reference

| PC1 | PC2 | Description |
|-----|-----|-------------|
| IS | ETW | Income Support — Expected to Work |
| IS | BFE | Income Support — Barriers to Full Employment |
| OTI | OTI | One Time Issues |
| AISH | AISH | Assured Income for Severely Handicapped |

For the full taxonomy with client type codes, see `ai/project/glossary.md`.

# CACHE Manifest Summary

For complete documentation of data coverage, quality metrics, and variable availability, see `data-public/metadata/CACHE-manifest.md`.

Key coverage facts:
- **CLIENT_ROSTER**: ~1.35M persons, core identifiers only
- **SUPPORT_BY_YEAR**: ~1.85M person-year records, 354K unique individuals, demographics 50.5% coverage
- **TRAINING_BY_YEAR**: ~525K person-year records, 268K individuals, demographics 99.7% coverage
- **ASSESSMENT_BY_YEAR**: Assessment events (coverage varies by type)
